---
title: "Reanalysis of Scheibehenne, Jamil, & Wagenmakers (2016)"
subtitle: "Hotel towel reuse"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

# Overview

- Computes joint BFs for the set of experiments reported in from Scheibehenne et al. (2016).
- Source paper: Scheibehenne, B., Jamil, T., & Wagenmakers, E.-J. (2016). Bayesian Evidence Synthesis Can Reconcile Seemingly Inconsistent Results: The Case of Hotel Towel Reuse. Psychological Science, 27(7), 1043-1046.
- Each experiment is analyzed as one 2x2 contingency table of counts:
  - rows: condition (control, social norm)
  - columns: towel reuse outcome (yes, no)
- For each study, evidence is computed for:
  - predicted hypothesis: social norm > control (higher reuse probability)
  - unconstrained hypothesis
  - complement: social norm < control

```{r packages, message=FALSE}
library(BayesFactor)
library(dplyr)
library(here)

source(here("functions/compute_joint_complements.r"))

# helper functions for formatting numbers
fmt2 <- function(x) sprintf("%.2f", x)
fmt4 <- function(x) sprintf("%.4f", x)
fmtE <- function(x) sprintf("%.2e", x)
```

# Data

```{r data}
rawData <- read.table(
  here("data/towelData.csv"),
  header = TRUE,
  sep = ";",
  stringsAsFactors = FALSE
)

# n experiments, each represented by 4 rows (from 2x2 contingency table)
n <- nrow(rawData) / 4

# Add pooled "Total" 2x2 table as in the original script
bigTable <- array(rawData$Count, dim = c(2, 2, n))
total2x2 <- apply(bigTable, 2, rowSums)

totalFrame <- data.frame(
  Source = rep("Total", 4),
  AuthorName = rep("Total", 4),
  Experiment = rep(NA, 4),
  Year = rep(NA, 4),
  Group = c("Control", "Social Norm", "Control", "Social Norm"),
  Towel.Reuse = c("Yes", "Yes", "No", "No"),
  Count = c(total2x2[1, 1], total2x2[2, 1], total2x2[1, 2], total2x2[2, 2])
)

rawData <- rbind(rawData, totalFrame)
```

# Study-specific Bayes factors

```{r study-bfs}
# Names of experiments
exList <- unique(rawData$Source)

# Initialize vectors to store study-specific Bayes factors
bf_1u_vec <- rep(NA_real_, length(exList))
bf_cu_vec <- rep(NA_real_, length(exList))
bf_10_vec <- rep(NA_real_, length(exList))
bf_1c_vec <- rep(NA_real_, length(exList))

# Set seed for reproducibility
set.seed(45)

# Compute BFs for each experiment
for (i in seq_along(exList)) {

  # Extract the 2x2 contingency table for the current experiment:
  # rows = condition (control, social norm), columns = towel reuse (yes, no)
  thisEx <- rawData[rawData$Source == exList[i], ]
  table2x2 <- cbind(yes = thisEx[1:2, "Count"], no = thisEx[3:4, "Count"])
  rownames(table2x2) <- c("control", "socialNorm")

  # Fit the unconstrained model H_u against the null model H_0:
  # - sampleType = "indepMulti", fixedMargin = "rows": the number of guests per
  #   condition is fixed, so each condition is an independent binomial with its
  #   own reuse proportion (theta_c for control, theta_e for social norm)
  # - priorConcentration = 1: independent uniform Beta(1, 1) priors on theta_c
  #   and theta_e (as in Scheibehenne et al., 2016, p. 1043)
  # - H_0 states that both conditions share the same proportion
  X <- contingencyTableBF(
    table2x2,
    sampleType = "indepMulti",
    fixedMargin = "rows",
    priorConcentration = 1
  )

  # Unused draws, kept from the original script: they consume random numbers,
  # so removing them would change the posterior samples below (and thus the
  # reported Bayes factors)
  post_alt  <- posterior(X, iterations = 10000)
  post_null <- posterior(1 / X, iterations = 10000)

  # BF_u0: evidence for H_u against H_0 (computed analytically by BayesFactor;
  # X stores it on the log scale)
  bf_u0 <- exp(as.numeric(X@bayesFactor$bf))

  # Draw 10,000 posterior samples of the two proportions under H_u
  # (these follow Beta(1 + yes, 1 + no) distributions in each condition)
  post.result <- BayesFactor::posterior(X, iterations = 10000)
  idx <- grep("omega", colnames(post.result))
  theta <- as.data.frame(post.result[, idx])
  # column 1: theta_c, proportion of towel reuse in the control condition
  # column 2: theta_e, proportion of towel reuse in the social-norm condition

  # Fit of H_1 (theta_e > theta_c) and of the complement H_c (theta_e < theta_c):
  # the proportions of posterior samples in agreement with each hypothesis
  prop.consistent <- mean(theta[, 1] < theta[, 2])
  prop.complement <- mean(theta[, 1] > theta[, 2])

  # BF_1u = fit / complexity; the complexity is P(theta_e > theta_c) under the
  # uniform priors, which is 0.5 because the prior is symmetric
  bf_1u <- prop.consistent / 0.5

  # BF_10 by transitivity: BF_10 = BF_u0 * BF_1u
  bf_10 <- bf_u0 * bf_1u

  # BF_cu for the complement, with complexity 1 - 0.5 = 0.5;
  # via BF_c0 = BF_u0 * BF_cu, so BF_cu = BF_c0 / BF_u0 = prop.complement / 0.5
  bf_c0 <- bf_u0 * prop.complement / 0.5
  bf_cu <- bf_c0 / bf_u0

  # BF_1c by transitivity: BF_1c = BF_1u / BF_cu
  bf_1c <- bf_1u / bf_cu

  # Store the study-specific BFs
  bf_1u_vec[i] <- bf_1u
  bf_cu_vec[i] <- bf_cu
  bf_10_vec[i] <- bf_10
  bf_1c_vec[i] <- bf_1c
}

# Create data frame with results
results_df <- data.frame(
  Study = exList,
  bf_10 = bf_10_vec,
  bf_1u = bf_1u_vec,
  bf_1c = bf_1c_vec,
  bf_cu = bf_cu_vec
) %>%
  filter(Study != "Total")

# Create dataframe with only the relevant columns for the compute_joint_complements function
final_df <- results_df %>%
  select(Study, bf_1u, bf_cu) %>%
  mutate(P_theta_in_Hc = 0.5)

# T = nr of experiments
T <- nrow(final_df)
```

```{r study-bfs-table, echo=FALSE}
knitr::kable(
  results_df,
  digits = 3,
  col.names = c(
    "Study",
    "$\\mathrm{BF}_{1,0}^{(t)}$",
    "$\\mathrm{BF}_{1,u}^{(t)}$",
    "$\\mathrm{BF}_{1,c}^{(t)}$",
    "$\\mathrm{BF}_{c,u}^{(t)}$"
  ),
  escape = FALSE
)
```

# Product Bayes factors

```{r product-bfs}
bf_10 <- prod(results_df$bf_10)
bf_1u <- prod(results_df$bf_1u)
bf_1c <- prod(results_df$bf_1c)
```

- $\mathrm{BF}_{1,0}^{(\mathrm{joint})} = `r fmtE(bf_10)`$.
- $\mathrm{BF}_{1,u}^{(\mathrm{joint})} = `r fmt4(bf_1u)`$.
- $\mathrm{BF}_{1,c}^{(\mathrm{joint})} = `r fmtE(bf_1c)`$.

# BF against complete joint complement

## 1: Enumeration (Eq. 7)

```{r enumeration}
set.seed(45)
res <- compute_joint_complements(
  final_df,
  replace_0_with = 0.001,
  round_res = FALSE
)

# Sanity check: BF_1u(joint) / BF_cu(joint) should equal the product of the BF_1c's
all.equal(res["BES_c", "bf_1u"] / res["BES_c", "bf_cu"], bf_1c)

bf_1c_star_enumeration <- res["BES_c", "bf_1u"] / res["BES_c_star", "bf_cu_star"]
```

## 2: Complexity weighting (Eq. 8)

- Check: with $c_1^{(t)} = 0.5$ for all studies, this should be close to equal weighting.
- Uses $\mathrm{BF}_{1,u}^{(t)}$ and complexities $c_1^{(t)}$.

$$
\mathrm{BF}^{(\text{joint})}_{1,c^*}
= \frac{\prod_{t=1}^{T}\mathrm{BF}^{(t)}_{1,u}\,\bigl(1-\prod_{t=1}^{T} c_1^{(t)}\bigr)}
  {1-\prod_{t=1}^{T} c_1^{(t)}\,\mathrm{BF}^{(t)}_{1,u}}
$$

```{r complexity-weighting}
c1 <- 1 - final_df$P_theta_in_Hc

bf_1c_star_formula_complexity <-
  prod(results_df$bf_1u) * (1 - prod(c1)) / (1 - prod(c1 * results_df$bf_1u))
```

## 3: Equal weighting (Eq. 9)

- Uses only $\mathrm{BF}_{1,c}^{(t)}$.

$$
\mathrm{BF}^{(\text{joint})}_{1,c^*}
= \frac{2^T-1}{\prod_{t=1}^{T}\bigl(1 + 1/\mathrm{BF}^{(t)}_{1,c}\bigr) - 1}
$$

```{r equal-weighting}
bf_1c_star_formula_equal <- (2^T - 1) / (prod(1 + 1 / results_df$bf_1c) - 1)
```

# Summary

```{r summary-table, echo=FALSE}
knitr::kable(
  data.frame(
    Metric = c(
      "$\\mathrm{BF}_{1,0}^{(\\mathrm{joint})}$",
      "$\\mathrm{BF}_{1,u}^{(\\mathrm{joint})}$",
      "$\\mathrm{BF}_{1,c}^{(\\mathrm{joint})}$",
      "$\\mathrm{BF}_{1,c^*}^{(\\mathrm{joint})}$ (Eq. 7)",
      "$\\mathrm{BF}_{1,c^*}^{(\\mathrm{joint})}$ (Eq. 8)",
      "$\\mathrm{BF}_{1,c^*}^{(\\mathrm{joint})}$ (Eq. 9)"
    ),
    Value = c(
      formatC(bf_10, format = "e", digits = 3),
      formatC(bf_1u, format = "f", digits = 4),
      formatC(bf_1c, format = "e", digits = 3),
      formatC(bf_1c_star_enumeration, format = "f", digits = 4),
      formatC(bf_1c_star_formula_complexity, format = "f", digits = 4),
      formatC(bf_1c_star_formula_equal, format = "f", digits = 4)
    ),
    check.names = FALSE
  ),
  col.names = c("", ""),
  align = c("l", "r"),
  escape = FALSE
)
```

# Session info

```{r session-info}
sessionInfo()
```