---
title: "Reanalysis of Volker (2022)"
subtitle: "Network embeddedness, trust, and cooperation"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

# Overview

- Reproduces Table 2 of the paper.
- Source: Volker, T. B. (2022). *The future is made today: Concerns for reputation
  foster trust and cooperation* [Unpublished master's thesis]. Utrecht University.
  [PDF](https://github.com/thomvolker/bes_master_thesis_sasr/blob/main/thesis/thesis_volker.pdf)
- 16 study-specific tests from 8 studies, each on a regression coefficient $\beta$:
    - $H_1: \beta > 0$
    - $H_u: \beta \in \mathbb{R}$
    - $H_c: \beta \le 0$
- Complexity $c_1^{(t)} = 0.5$ in all tests: the prior
  $\mathrm{N}\left(0, \hat{\Sigma}_{\beta} / b\right)$ is centered at 0 (Volker, 2022, pp. 32, 36).

```{r packages, message=FALSE}
library(dplyr)
library(here)

# Function that enumerates all components of the complete joint complement
source(here("functions/compute_joint_complements.r"))

# Helpers to format numbers in the text
fmt2 <- function(x) sprintf("%.2f", x)  # 2 decimals
fmt4 <- function(x) sprintf("%.4f", x)  # 4 decimals
fmtE <- function(x) {                   # scientific notation for LaTeX math
    if (!is.finite(x)) return("\\infty")
    e <- floor(log10(abs(x)))
    sprintf("%.2f \\times 10^{%d}", x / 10^e, e)
}
```

# Data

- Layout from Volker's Table 2; Bayes factors from his Table 4 (3 decimals).
- Trustfulness: proportion of the endowment sent to the trustee.
- Trustworthiness: proportion of the amount sent that is returned (p. 27).

```{r data}
table2_df <- data.frame(
    study = c(
        rep("bolton_2004", 2),
        rep("duffy_2013_no_netmin", 2),
        rep("duffy_2013_no_netfull", 2),
        rep("buskens_2010", 2),
        rep("van_miltenburg_2012", 2),
        rep("frey_2019", 2),
        rep("barrera_buskens_2009", 2),
        "seinen_schram_2006",
        "corten_2016"
    ),
    outcome = c(
        rep(c("Trustfulness", "Trustworthiness"), 7),
        "Cooperation (helping)",
        "Cooperation"
    ),
    no_network = c(
        0.67, 0.69, 0.62, 0.68, 0.75, 0.62, 0.83, 0.87,
        0.90, 0.90, 0.77, 0.67, 0.73, 0.46, 0.38, 0.49
    ),
    network = c(
        0.75, 0.61, 0.64, 0.77, 0.86, 0.94, 0.94, 0.94,
        0.88, 0.91, 0.81, 0.88, 0.70, 0.50, 0.61, 0.37
    ),
    bf_iu = c(
        1.474, 0.642, 1.292, 1.823, 1.799, 2.000, 1.848,
        1.680, 0.675, 1.115, 1.426, 1.998, 0.482, 1.450,
        1.902, 0.102
    ),
    bf_ic = c(
        2.799, 0.473, 1.825, 10.319, 8.958, 2.60e+04,
        12.168, 5.255, 0.509, 1.260, 2.486, 1.01e+03,
        0.318, 2.634, 19.351, 0.054
    )
)
```

- This is a check: in the special case where $c_1^{(t)} = 0.5$ for all studies,
the result should equal the one from equal weighting.
- Volker reports $\mathrm{BF}_{1,u}^{(t)}$ and $\mathrm{BF}_{1,c}^{(t)}$.
- `compute_joint_complements()` also needs $\mathrm{BF}_{c,u}^{(t)}$, which can be obtained by
  transitivity: $\mathrm{BF}_{c,u}^{(t)} = \mathrm{BF}_{1,u}^{(t)} / \mathrm{BF}_{1,c}^{(t)}$.

```{r study-bfs}
results_df <- table2_df %>%
    transmute(
        Study   = study,
        Outcome = outcome,
        bf_1u   = bf_iu,
        bf_1c   = bf_ic,
        bf_cu   = bf_iu / bf_ic
    )

# Input for compute_joint_complements():
# P_theta_in_Hc is the complexity of the complement, 1 - c_1 = 0.5
table2_ready <- results_df %>%
    select(Study, bf_1u, bf_cu) %>%
    mutate(P_theta_in_Hc = 0.5)

T <- nrow(results_df)  # number of study-specific tests (16)
```

```{r study-bfs-table, echo=FALSE}
knitr::kable(
    results_df,
    digits = 3,
  col.names = c("Study", "Outcome",
          "$\\mathrm{BF}_{1,u}$",
          "$\\mathrm{BF}_{1,c}$",
          "$\\mathrm{BF}_{c,u}$"),
  escape = FALSE
)
```

# Product Bayes factors

```{r product-bfs}
bf_1u <- prod(results_df$bf_1u)  # against the joint unconstrained hypothesis
bf_1c <- prod(results_df$bf_1c)  # against the incomplete joint complement
```

- $\mathrm{BF}_{1,u}^{(\text{joint})} = `r fmt2(bf_1u)`$ (Volker: 7.27; difference due
  to rounding).
- $\mathrm{BF}_{1,c}^{(\text{joint})} = `r fmtE(bf_1c)`$.

# BF against the complete joint complement

- Complete joint complement: $H_1^{(t)}$ is false in at least one test.
- Computed in three ways; with exact inputs, all three agree here (with
  $c_1^{(t)} = 0.5$, equal and complexity weighting coincide).

## 1: Enumeration (Eq. 7)

- Sums over all $2^T - 1 = `r format(2^T - 1, big.mark = ",")`$ components with equal
  weights, giving $\mathrm{BF}_{c^*,u}^{(\text{joint})}$.
- Then $\mathrm{BF}_{1,c^*}^{(\text{joint})} = \mathrm{BF}_{1,u}^{(\text{joint})} / \mathrm{BF}_{c^*,u}^{(\text{joint})}$.

```{r enumeration}
res <- compute_joint_complements(table2_ready,
                                 replace_0_with = 0.001,
                                 round_res = FALSE)

# Sanity check: BF_1u(joint) / BF_cu(joint) should equal the product of the BF_1c's
all.equal(res["BES_c", "bf_1u"] / res["BES_c", "bf_cu"], bf_1c)

bf_1c_star_enumeration <- res["BES_c", "bf_1u"] / res["BES_c_star", "bf_cu_star"]
```

## 2: Complexity weighting (Eq. 8)

- Check: in the special case where $c_1^{(t)} = 0.5$ for all studies,
  the result should equal the one obtained from equal weighting.
- Uses the $\mathrm{BF}_{1,u}^{(t)}$ and complexities $c_1^{(t)}$.

$$
\mathrm{BF}^{(\text{joint})}_{1,c^*}
= \frac{\prod_{t=1}^{T}\mathrm{BF}^{(t)}_{1,u}\,\bigl(1-\prod_{t=1}^{T} c_1^{(t)}\bigr)}
       {1-\prod_{t=1}^{T} c_1^{(t)}\,\mathrm{BF}^{(t)}_{1,u}}
$$

```{r complexity-weighting}
c1 <- 1 - table2_ready$P_theta_in_Hc  # complexity c_1^(t) of H_1 in each test

bf_1c_star_formula_complexity <-
    prod(results_df$bf_1u) * (1 - prod(c1)) / (1 - prod(c1 * results_df$bf_1u))
```

## 3: Equal weighting (Eq. 9)

- Uses only the $\mathrm{BF}_{1,c}^{(t)}$.

$$
\mathrm{BF}^{(\text{joint})}_{1,c^*}
= \frac{2^T-1}{\prod_{t=1}^{T}\bigl(1 + 1/\mathrm{BF}^{(t)}_{1,c}\bigr) - 1}
$$

```{r equal-weighting}
bf_1c_star_formula_equal <- (2^T - 1) / (prod(1 + 1 / results_df$bf_1c) - 1)
```

## Comparison

```{r comparison, echo=FALSE}
knitr::kable(
    data.frame(
        Method = c("Enumeration (Eq. 7)",
             "Complexity weighting (Eq. 8)",
             "Equal weighting (Eq. 9)"),
      `Computed from` = c("$\\mathrm{BF}_{1,u}^{(t)}$",
        "$\\mathrm{BF}_{1,u}^{(t)}$ and $c_1^{(t)}$",
        "$\\mathrm{BF}_{1,c}^{(t)}$"),
        `BF_1c*(joint)` = c(bf_1c_star_enumeration,
                bf_1c_star_formula_complexity,
                bf_1c_star_formula_equal),
        check.names = FALSE
    ),
    col.names = c("Method",
                  "Computed from",
                  "$\\mathrm{BF}_{1,c^*}^{(\\mathrm{joint})}$"),
    digits = 4,
    escape = FALSE
)
```

- Enumeration and Eq. 8: `r fmt2(bf_1c_star_enumeration)`; Eq. 9:
  `r fmt2(bf_1c_star_formula_equal)`.
- The difference comes from *which* rounded Bayes factors each method uses:
    - Enumeration: numerator $\prod_t \mathrm{BF}_{1,u}^{(t)}$; denominator
      $\mathrm{BF}_{c^*,u}^{(\text{joint})} \approx 1$.
    - Eq. 8: $\mathrm{BF}_{1,u}^{(t)}$.
    - Eq. 9: $\mathrm{BF}_{1,c}^{(t)}$.

# Rounding check

- **Goal:** How much of the gap between
  $\mathrm{BF}_{1,u}^{(\text{joint})} = `r fmt2(bf_1u)`$ and
  $\mathrm{BF}_{1,c^*}^{(\text{joint})} = `r fmt2(bf_1c_star_formula_equal)`$ (Eq. 9)
  is due to rounding?
- The two are computed from different columns
  ($\mathrm{BF}_{1,u}$ vs. $\mathrm{BF}_{1,c}$), which Volker rounded separately to 3
  decimals.
- **Logic:** With $c_1 = 0.5$, each $\mathrm{BF}_{1,u}^{(t)}$ determines its
  $\mathrm{BF}_{1,c}^{(t)}$ exactly, and vice versa:
  $$
  \mathrm{BF}_{1,c}^{(t)} = \frac{\mathrm{BF}_{1,u}^{(t)}}{2 - \mathrm{BF}_{1,u}^{(t)}}
  \qquad\Longleftrightarrow\qquad
  \mathrm{BF}_{1,u}^{(t)} = \frac{2\,\mathrm{BF}_{1,c}^{(t)}}{1 + \mathrm{BF}_{1,c}^{(t)}}
  $$
    - Set 1: reported $\mathrm{BF}_{1,u}^{(t)}$ → implied $\mathrm{BF}_{1,c}^{(t)}$.
    - Set 2: reported $\mathrm{BF}_{1,c}^{(t)}$ → implied $\mathrm{BF}_{1,u}^{(t)}$.
    - Within a set, both joint BFs use the same information, so any difference is
      genuine (not due to rounding).
- **Expectation**

- With $c_1^{(t)} = 0.5$, Eqs. 8 and 9 reduce to
  $\mathrm{BF}^{(\text{joint})}_{1,c^*} = \mathrm{BF}^{(\text{joint})}_{1,u} \times \text{factor}$,
  where $\text{factor} = (2^T-1)/(2^T-\mathrm{BF}^{(\text{joint})}_{1,u})$.

```{r expectation}
# Step 1: ratio BF_1c*(joint) / BF_1u(joint) implied by Eqs. 8 and 9 when c_1 = 0.5
factor <- (2^T - 1) / (2^T - bf_1u)

# Step 2: expected BF_1c*(joint) without rounding
bf_1cstar_exp <- bf_1u * factor

# Step 3: expected difference between BF_1c*(joint) and BF_1u(joint)
diff_exp <- bf_1cstar_exp - bf_1u
```

With $T = `r T`$ and $\mathrm{BF}^{(\text{joint})}_{1,u} = `r sprintf("%.6f", bf_1u)`$:

1. **Factor:** how much larger $\mathrm{BF}^{(\text{joint})}_{1,c^*}$ should be than
   $\mathrm{BF}^{(\text{joint})}_{1,u}$ (here `r sprintf("%.4f", 100 * (factor - 1))`% larger):
   $$
   \text{factor} = \frac{`r sprintf("%.0f", 2^T - 1)`}{`r sprintf("%.0f", 2^T)` - `r sprintf("%.6f", bf_1u)`} = `r sprintf("%.9f", factor)`
   $$
2. **Expected $\mathrm{BF}^{(\text{joint})}_{1,c^*}$** without rounding:
   $$
   `r sprintf("%.6f", bf_1u)` \times `r sprintf("%.9f", factor)` = `r sprintf("%.6f", bf_1cstar_exp)`
   $$
3. **Expected within-set difference:**
   $$
   `r sprintf("%.6f", bf_1cstar_exp)` - `r sprintf("%.6f", bf_1u)` = `r sprintf("%.6f", diff_exp)` \approx `r fmt4(diff_exp)`
   $$

- This is the genuine difference that remains without rounding. The table below
  compares it with the observed differences.

```{r rounding-check}
# Set 1: consistent set based on the reported BF_1u's
bf_1c_from_1u <- results_df$bf_1u / (2 - results_df$bf_1u)

check_set1_bf_1u_joint <- prod(results_df$bf_1u)
check_set1_bf_1cstar   <- (2^T - 1) / (prod(1 + 1 / bf_1c_from_1u) - 1)

# Set 2: consistent set based on the reported BF_1c's
bf_1u_from_1c <- 2 * results_df$bf_1c / (1 + results_df$bf_1c)

check_set2_bf_1u_joint <- prod(bf_1u_from_1c)
check_set2_bf_1cstar   <- (2^T - 1) / (prod(1 + 1 / results_df$bf_1c) - 1)
```

```{r rounding-check-table, echo=FALSE}
knitr::kable(
    data.frame(
    Inputs = c("Set 1: reported $\\mathrm{BF}_{1,u}$ values",
           "Set 2: reported $\\mathrm{BF}_{1,c}$ values",
                   "Reported values (mixed sets)"),
        `BF_1u(joint)` = c(check_set1_bf_1u_joint, check_set2_bf_1u_joint, bf_1u),
        `BF_1c*(joint)` = c(check_set1_bf_1cstar, check_set2_bf_1cstar,
                            bf_1c_star_formula_equal),
        check.names = FALSE
    ) %>%
        mutate(Difference = `BF_1c*(joint)` - `BF_1u(joint)`),
  col.names = c("Inputs",
          "$\\mathrm{BF}_{1,u}^{(\\mathrm{joint})}$",
          "$\\mathrm{BF}_{1,c^*}^{(\\mathrm{joint})}$",
          "Difference"),
  digits = 4,
  escape = FALSE
)
```

**Conclusion:**

- Within each set: difference $\approx$ `r fmt4(check_set1_bf_1cstar - check_set1_bf_1u_joint)`,
  as expected.
- Reported (mixed) values: difference = `r fmt4(bf_1c_star_formula_equal - bf_1u)`.
- So the gap is almost entirely due to rounding, mostly from the smallest Bayes
  factors (e.g., Corten et al., 2016: reported $\mathrm{BF}_{1,u} = 0.102$, but
  $\mathrm{BF}_{1,c} = 0.054$ implies
  $\mathrm{BF}_{1,u} = `r fmt4(bf_1u_from_1c[results_df$Study == "corten_2016"])`$).

# Summary

```{r summary-table, echo=FALSE}
knitr::kable(
    data.frame(
    Metric = c("$\\mathrm{BF}_{1,u}^{(\\mathrm{joint})}$",
           "$\\mathrm{BF}_{1,c}^{(\\mathrm{joint})}$",
           "$\\mathrm{BF}_{1,c^*}^{(\\mathrm{joint})}$"),
    Value = formatC(c(bf_1u, bf_1c, bf_1c_star_formula_equal),
            digits = 3, format = "g"),
        check.names = FALSE
    ),
  col.names = c("", ""),
  align = c("l", "r"),
  escape = FALSE
)
```

# Session info

```{r session-info}
sessionInfo()
```
